Generalized Derivation for the Motion of an n-Pendulum System

### 2021-05-22 paper/paper.tex
add diagram

% @Author: UnsignedByte
% @Date:   11:12:32, 08-Dec-2020
% @Last Modified by:   UnsignedByte
% @Last Modified time: 2021-05-11 13:44:56

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\usepackage[%  
    colorlinks=true,
    pdfborder={0 0 0},
    linkcolor=black
]{hyperref}
\title{Generalized Derivation for the Motion of an n-Pendulum System}
\author{Edmund Lam \and Nicolas Wieczorek}
\begin{document}
\begin{titlepage}
	\maketitle
	\thispagestyle{empty}
\end{titlepage}
\newpage
  \tableofcontents
\newpage

@@

\section{Lagrangian Approach}
In contrast to the limitations of the Newtonian Free Body Diagram approach to calculating propogated forces amongst an $n$-pendulum system, the use of Lagrangian Mechanics does not require an analysis of forces. Instead, the total potential and kinetic energy in the system is used to determine equations of motion for each pendulum in the system.
\subsection{Derivation of Equations of Motion for Finite \texorpdfstring{$n$}{n}}
The use of the lagrangian method requires on two main parts: defining constraints and variables, and using them to create equations of motion.
\subsubsection{Deriving the Lagrangian \texorpdfstring{$L=T-V$}{L=T-V}}
We begin by defining the kinematic constraints of our problem.
\begin{align*}
	x_i &= x_{i-1} + l\sin{\theta_i}\\
	&= l\sum_{k=1}^i\sin{\theta_k}\\
	y_i &= y_{i-1} - l\cos{\theta_i}\\
	&= -l\sum_{k=1}^i\cos{\theta_k}
\end{align*}
Note that $y_i$ is negative, because $\theta_i$ is to the negative vertical. Next we derive the velocities. These can be calculated simply by taking the derivative of the positions.
\begin{align*}
	\dot{x}_i &= \dot x_{i-1} + \dot\theta_i \cdot l\cos{\theta_i}\\
	&= l\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\\
	\dot{x}_i &= \dot y_{i-1} + \dot\theta_i \cdot l\sin{\theta_i}\\
	&= l\sum_{k=1}^i\dot\theta_k\sin{\theta_k}
\end{align*}
Plugging these in, we can calculate the total potential energy $V=mgh$ and the total kinetic energy $T=\frac{1}{2}mv^2$. First, we solve for $V$ and simplify:
\begin{align*}
	V &= mg\sum_{i=1}^n y_i \\
	&=-mgl\sum_{i=1}^n\sum_{k=1}^i\cos\theta_k
	&\text{(Substitute in $y_i$)}\\
	&=-mgl\sum_{i=1}^n (n-i+1)\cos\theta_i
	&\text{(Each element $\cos\theta_i$ appears $(n-i+1)$ times)}\\
\end{align*}
And we can do the same with $T$, substituting in $v^2=\sum_{i=1}^n\left(\dot{x}_i^2+\dot{y}_i^2\right)$.
\begin{align*}
	T &= \frac{m}{2}\sum_{i=1}^n \left(\dot{x}_i^2+\dot{y}_i^2\right)\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \left(\left(\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\right)^2+\left(\sum_{k=1}^i\dot\theta_k\sin{\theta_k}\right)^2\right)
	&\text{(Substitute in $\dot x_i$ and $\dot y_i$)}\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \sum_{j=1}^i\sum_{k=1}^i\dot\theta_j\dot\theta_k\left(\cos{\theta_j}\cos{\theta_k}+\sin{\theta_j}\sin{\theta_k}\right)
	&\text{(Expand squared sums)}\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \sum_{j=1}^i\sum_{k=1}^i\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)
	&\text{(Simplify using $\cos$ addition formula)}\\
	&= \frac{ml^2}{2} \sum_{j=1}^n\sum_{k=1}^n(n-\max(j,k)+1)\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)
	&\text{(Simplify out first sum)}\\
\end{align*}
Next, we can plug $T$ and $V$ into the lagrangian, giving us the following:
\begin{align*}
	L &= T-V \\
	&=  \frac{ml^2}{2} \sum_{j=1}^n\sum_{k=1}^n(n-\max(j,k)+1)\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k) + mgl\sum_{i=1}^n (n-i+1)\cos\theta_i\\
\end{align*}
\subsubsection{Utilizing the Euler-Lagrange Equation}
Now that we have our $L$ expression, we can use this to solve for the angular acceleration of each pendulum using the Euler-Lagrange Equation. We start with the equation $0 = \frac{d}{dt}\left(\frac{\partial L}{\partial \dot \theta_i}\right) - \frac{\partial L}{\partial \theta_i}$, and we will use this equation to solve for all $\ddot\theta_i.$

@@

Once again, we start by solving for $\frac{\partial L}{\partial \dot \theta}$ by taking a derivative of $L$ with respect to $\dot\theta$. We can use the same tricks as before, but this time the $i=i$ case does not need to be specially considered, as the integral is continuous and a single individual value makes an insignificant change to the whole.
\begin{align*}
&mL^2\frac{\partial L}{\partial \dot \theta_i}\left(\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)dk\right)\\
&= mL^2\int_0^L(L-\max(i,k))\dot\theta_k\cos(\theta_i-\theta_k)dk
\end{align*}
Again, we take the derivative with respect to time:
\begin{align*}
&= mL^2\frac{d}{dt}\int_0^L(L-\max(i,k))\dot\theta_k\cos(\theta_i-\theta_k)dk\\
&= mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)dk
\end{align*}
With $\frac{d}{dt}\left(\frac{\partial L}{\partial \dot\theta_i}\right)$ finished, we once again calculate the remaining term $\frac{\partial L}{\partial \theta_i}$.
\begin{align*}
&mL^2\frac{\partial L}{\partial \theta_i}\left(\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)dk\right) + \frac{\partial L}{\partial \theta_i}mgL(L-i)\cos\theta_i\\
&= mL^2\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)dk - mgL(L-i)\sin\theta_i\\
\end{align*}
Now, we combine this back into the Euler-Lagrange equation, and simplify.
\begin{align*}
0&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)dk\\
&\qquad\qquad+ mL^2\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)dk + mgL(L-i)\sin\theta_i\\
&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k) + \dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)dk\\
&\qquad\qquad+ mgL(L-i)\sin\theta_i\\
&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)+\dot\theta_k^2\sin(\theta_i-\theta_k)\right)dk+ mgL(L-i)\sin\theta_i\\
\end{align*}
We have our equation 
$$0=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)+\dot\theta_k^2\sin(\theta_i-\theta_k)\right)dk+ mgL(L-i)\sin\theta_i$$
for all $i \in [0,L]$, which we next need to evaluate for all $\ddot\theta_i$.
\subsubsection{Finding \texorpdfstring{$\ddot\theta_i$}{Angular Acceleration}}
e
\end{document}

>>>>

% @Author: UnsignedByte
% @Date:   11:12:32, 08-Dec-2020
% @Last Modified by:   UnsignedByte
% @Last Modified time: 2021-05-21 22:45:03

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{asymptote}
\graphicspath{ {./images/} }
\usepackage[%  
    colorlinks=true,
    pdfborder={0 0 0},
    linkcolor=black
]{hyperref}
\title{Generalized Derivation for the Motion of an n-Pendulum System}
\author{Edmund Lam \and Nicolas Wieczorek}
\begin{document}
\begin{titlepage}
	\maketitle
	\thispagestyle{empty}
\end{titlepage}
\newpage
  \tableofcontents
\newpage

\begin{asydef}
import astylib;
\end{asydef}

@@

A single pendulum, is defined by a point of mass suspended by a rigid rod allowed to rotate freely on a joint at the other end of the rod. The mechanics of such a pendulum is simple to calculate, even with Newtonian mechanics. When attaching a second pendulum to the mass of the first pendulum, we get a system widely known as the "double pendulum." The double pendulum system is widely known to be a chaotic system, meaning a small difference in initial conditions results in widely different motion after a period of time. Figure \ref{fig:1} is an example of such a system. This method can be continued, adding more and more pendulums, creating a "pendulum chain"--an N pendulum system.

To simplify the problem, we define our system while leaving a number of things constant. Our pendulums will have massless rods of constant length $l$, with a mass $m$ at the end of each rod, or at the "joint" between two pendulums. Each pendulum will be defined by a single angle $\theta_i$, the angle of the pendulum rod from the vertical.

\begin{figure}[h]
	\caption{Example double pendulum system with labeled components.}
	\label{fig:1}
	\begin{center}
		\begin{asy}
		pair R,A,B,AY,BY;
		real TA,TB,L;
		L = 100; // length
		astyscale = 15;

		R = (0,0);
		TA = 45;
		A = R+dir(270+TA)*L;
		AY = (R.x, A.y);
		TB = -30;
		B = A+dir(270+TB)*L;
		BY = (A.x, B.y);

		pen guide = linetype(new real[] {0,5});

		draw(R--A--B);
		draw(astyangle(AY, R, A));
		draw(astyangle(B, A, BY));
		draw(astycomponents(R, A), guide);
		draw(astycomponents(A, B), guide);

		label("$R$", R, -dir(270+TA));
		label("$A$", A, E);
		label("$B$", B, dir(270+TB));
		label("$l$", R--A, dir(TA));
		label("$l$", A--B, -dir(TB));
		label("$y_1$", (R.x, A.y)--R, W);
		label("$x_1$", (R.x, A.y)--A, S);
		label("$y_2$", (A.x, B.y)--A, E);
		label("$x_2$", (A.x, B.y)--B, S);
		label("$\theta_1$", R+dir(270+TA/2)*astyscale, dir(270+TA/2));
		label("$\theta_2$", A+dir(270+TB/2)*astyscale, dir(270+TB/2));
		\end{asy}
	\end{center}
\end{figure}

\section{Lagrangian Approach}
In contrast to the limitations of the Newtonian Free Body Diagram approach to calculating propogated forces amongst an $n$-pendulum system, the use of Lagrangian Mechanics does not require an analysis of forces. Instead, the total potential and kinetic energy in the system is used to determine equations of motion for each pendulum in the system.
\subsection{Derivation of Equations of Motion for Finite \texorpdfstring{$n$}{n}}
The use of the lagrangian method requires on two main parts: defining constraints and variables, and using them to create equations of motion.
\subsubsection{Deriving the Lagrangian \texorpdfstring{$L=T-V$}{L=T-V}}
We begin by defining the kinematic constraints of our problem.
\begin{align*}
	x_i &= x_{i-1} + l\sin{\theta_i}\\
	&= l\sum_{k=1}^i\sin{\theta_k}\\
	y_i &= y_{i-1} - l\cos{\theta_i}\\
	&= -l\sum_{k=1}^i\cos{\theta_k}
\end{align*}
Note that $y_i$ is negative, because $\theta_i$ is to the negative vertical. Next we derive the velocities. These can be calculated simply by taking the derivative of the positions.
\begin{align*}
	\dot x_i &= \dot x_{i-1} + \dot\theta_i \cdot l\cos{\theta_i}\\
	&= l\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\\
	\dot y_i &= \dot y_{i-1} + \dot\theta_i \cdot l\sin{\theta_i}\\
	&= l\sum_{k=1}^i\dot\theta_k\sin{\theta_k}
\end{align*}
Plugging these in, we can calculate the total potential energy $V=mgh$ and the total kinetic energy $T=\frac{1}{2}mv^2$. First, we solve for $V$ and simplify:
\begin{align*}
	V &= mg\sum_{i=1}^n y_i \\
	&=-mgl\sum_{i=1}^n\sum_{k=1}^i\cos\theta_k
	&\text{(Substitute in $y_i$)}\\
	&=-mgl\sum_{i=1}^n (n-i+1)\cos\theta_i
	&\text{(Each element $\cos\theta_i$ appears $(n-i+1)$ times)}\\
\end{align*}
And we can do the same with $T$, substituting in $v^2=\sum_{i=1}^n\left(\dot{x}_i^2+\dot{y}_i^2\right)$.
\begin{align*}
	T &= \frac{m}{2}\sum_{i=1}^n \left(\dot{x}_i^2+\dot{y}_i^2\right)\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \left(\left(\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\right)^2+\left(\sum_{k=1}^i\dot\theta_k\sin{\theta_k}\right)^2\right)
	&\text{(Substitute in $\dot x_i$ and $\dot y_i$)}\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \sum_{j=1}^i\sum_{k=1}^i\dot\theta_j\dot\theta_k\left(\cos{\theta_j}\cos{\theta_k}+\sin{\theta_j}\sin{\theta_k}\right)
	&\text{(Expand squared sums)}\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \sum_{j=1}^i\sum_{k=1}^i\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)
	&\text{(Simplify using $\cos$ addition formula)}\\
	&= \frac{ml^2}{2} \sum_{j=1}^n\sum_{k=1}^n(n-\max(j,k)+1)\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)
	&\text{(Simplify out first sum)}\\
\end{align*}
Next, we can plug $T$ and $V$ into the lagrangian, giving us the following:
\begin{align*}
	L &= T-V \\
	&=  \frac{ml^2}{2} \sum_{j=1}^n\sum_{k=1}^n(n-\max(j,k)+1)\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k) + mgl\sum_{i=1}^n (n-i+1)\cos\theta_i\\
\end{align*}
\subsubsection{Utilizing the Euler-Lagrange Equation}
Now that we have our $L$ expression, we can use this to solve for the angular acceleration of each pendulum using the Euler-Lagrange Equation. We start with the equation $0 = \frac{d}{dt}\left(\frac{\partial L}{\partial \dot \theta_i}\right) - \frac{\partial L}{\partial \theta_i}$, and we will use this equation to solve for all $\ddot\theta_i.$

@@

Once again, we start by solving for $\frac{\partial L}{\partial \dot \theta}$ by taking a derivative of $L$ with respect to $\dot\theta$. We can use the same tricks as before, but this time the $i=i$ case does not need to be specially considered, as the integral is continuous and a single individual value makes an insignificant change to the whole.
\begin{align*}
&mL^2\frac{\partial L}{\partial \dot \theta_i}\left(\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)dk\right)\\
&= mL^2\int_0^L(L-\max(i,k))\dot\theta_k\cos(\theta_i-\theta_k)dk
\end{align*}
Again, we take the derivative with respect to time:
\begin{align*}
&= mL^2\frac{d}{dt}\int_0^L(L-\max(i,k))\dot\theta_k\cos(\theta_i-\theta_k)dk\\
&= mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)dk
\end{align*}
With $\frac{d}{dt}\left(\frac{\partial L}{\partial \dot\theta_i}\right)$ finished, we once again calculate the remaining term $\frac{\partial L}{\partial \theta_i}$.
\begin{align*}
&mL^2\frac{\partial L}{\partial \theta_i}\left(\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)dk\right) + \frac{\partial L}{\partial \theta_i}mgL(L-i)\cos\theta_i\\
&= mL^2\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)dk - mgL(L-i)\sin\theta_i\\
\end{align*}
Now, we combine this back into the Euler-Lagrange equation, and simplify.
\begin{align*}
0&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)dk\\
&\qquad\qquad+ mL^2\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)dk + mgL(L-i)\sin\theta_i\\
&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k) + \dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)dk\\
&\qquad\qquad+ mgL(L-i)\sin\theta_i\\
&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)+\dot\theta_k^2\sin(\theta_i-\theta_k)\right)dk+ mgL(L-i)\sin\theta_i\\
\end{align*}
We have our equation 
$$0=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)+\dot\theta_k^2\sin(\theta_i-\theta_k)\right)dk+ mgL(L-i)\sin\theta_i$$
for all $i \in [0,L]$, which we next need to evaluate for all $\ddot\theta_i$.
\end{document}
